---
title: "DA R PROGRAMMING EXAM"
author: "Perpetua odili"
date: "22/7/2024"
output: html_document
---

 ## IMPORT THE DATASET CALLED employees_attrition.csv file 
 ## INTO RSTUDIO
 
 ## EMPLOYEE ATTRITION MEANS EMPLOYEES ARE LEAVING THE COMPANY
 ## FOR ONE REASON OR THE OTHER.
 
 ## THE ATTRITION COLUMN HAS TWO POSSIBLE VALUES YES - FOR EMPLOYEES
 ## THAT LEFT THE COMPANY AND NO FOR THOSE THAT ARE  STILL AROUND.
 
 ## NOTE
 
 ## 1. RUN EVERY CODE CHUNCK YOU CREATE SO THAT THE OUTPUT WILL
 ## BE DISPLAYED BELOW THE CODE CHUNK.
 
 ## 2. SAVE YOUR RMARKDOWN FILE WITH YOUR NAME FOR IDENTIFICATION
 
 ## 3. USE DPLYR FUNCTIONS TO ANSWER THE FOLLOWING QUESTIONS.
 
 
 
```{r Q1}

# load the tidyverse package into this session

library("tidyverse")
```
 
```{r  Q2}
## import the given employees_attrition.csv file 
employees_attrition <-read.csv(file.choose())

```
```{r}
View(employees_attrition)
```

```{r Q3 }

## SHOW THE STRUCTURE OF THE DATAFRAME

str(employees_attrition)

```


```{r Q4 }

## HOW MANY ROWS AND COLUMNS ARE THERE IN THE DATAFRAME

num_rows <-nrow(employees_attrition)
print(num_rows)

num_cols <-ncol(employees_attrition)
print(num_cols)

```

```{r Q5}
## SHOW THE NUMBER OF EMPLOYEES THAT LEFT THE COMPANY AND THE 
## NUMBER THAT ARE STILL WITH THE COMPANY
## HINT - USE THE group_by() function to categorise.
# Load the necessary library
library(dplyr)


employees_attrition %>%
  group_by(attrition) %>%
  summarise(count = n())


```



```{r}
library(tidyverse)
```
```{r}
library(dplyr)
```



```{r}
```





```{r Q6}
## SHOW THE NUMBER OF EMPLOYEES THAT LEFT THE COMPANY IN EACH 
## DEPARTMENT.
employees_attrition %>%
  filter(attrition == "Yes")%>%
  group_by(attrition,department) %>%
  summarise(count = n())

```



```{r Q7}
## SHOW THE NUMBER OF MALE AND FEMALE EMPLOYEES THAT LEFT THE COMPANY
## AND THE NUMBER OF MALE AND FEMALE EMPLOYEES THAT ARE STILL WITH THE COMPANY
## HINT - USE THE group_by() function


library(dplyr)

employees_attrition %>%
  filter(attrition %in% c("Yes", "No")) %>%
  group_by(attrition, gender) %>%
  summarise(count = n(), .groups = 'drop')


```




```{r Q8}

## WHAT IS THE AVERAGE NUMBER OF YEARS WORKED IN THE COMPANY BY THOSE THAT LEFT THE COMPANY.
employees_attrition %>%
  filter(attrition == "Yes")%>%
  group_by(attrition) %>%
  summarise(year_sworked = mean(employees_attrition$YearsAtCompany))

```



```{r Q9}

```




```{r Q10}
## SHOW THE NUMBER OF EMPLOYEES THAT LEFT THE COMPANY BASED ON  THE isMarried column. 

employees_attrition %>%
  filter(attrition == "Yes")%>%
  group_by(isMarried) %>%
  summarise(count =n())


```



```{r Q11 }
## SHOW THE AVERAGE MONTHLY INCOME EARNED BY THOSE WHO LEFT 
## THE COMPANY. BREAK THE INFORMATION DOWN BY DEPARTMENTS
library(dplyr)


employees_attrition %>%
  filter(attrition == "Yes") %>%
  group_by(department) %>%
  summarise(avg_income = mean(monthly_income, na.rm = TRUE))



```


```{r Q12}

## SHOW THE AVERAGE MONTHLY INCOME EARNED BY THOSE WHO ARE STILL WITH THE COMPANY. BREAK THE INFORMATION DOWN BY DEPARTMENTS.
employees_attrition %>%
  filter(attrition == "No") %>%
  group_by(department) %>%
  summarise(avg_income = mean(monthly_income, na.rm = TRUE))

```



